---
title: "Untitled"
output: html_document
date: "2022-08-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
suppressMessages({
  library(tidyverse) #metapackage
  library(gt) #Easily Create Presentation-Ready Display Tables
  library(ggsci) #Scientific Journal and Sci-Fi Themed Color Palettes for 'ggplot2'
  library(reshape2) #flexibly Reshape Data
  library(ggrepel) #Automatically Position Non-Overlapping Text Labels with 'ggplot2'
  library(webshot)
  library(DescTools) # Tools for Descriptive Statistics
  library(ggimage) #geom layer for visualizing image files
  library(jpeg) #read and write jpeg files
  
  library(worldfootballR) #extract and clean world soccer data
  library(engsoccerdata) #English and European Soccer Results
  library(socceR) #Evaluating Sport Tournament Predictions
  library(ggsoccer) #Plot Soccer Event Data
  library(footBayes) #Fitting Bayesian and MLE Football Models
  library(itscalledsoccer) #American Soccer Analysis API Client
  library(FPLdata) #Read in Fantasy Premier League Data
  library(EUfootball) #Football Match Data of European Leagues
})
```

Lets check out the **worldfootballr** package

The statistic type options (stat_type) include:
"standard", "shooting", "passing", "passing_types", "gca", "defense", "possession", "playing_time", "misc", "keepers", "keepers_adv"
```{r}
fb_big5_player <- fb_big5_advanced_season_stats(
  season_end_year = 2023,
  stat_type = "standard",
  team_or_player = "player",
  time_pause = 3
)
```

Sort player data
```{r}
fb <- fb_big5_player %>% 
  select(Squad, Comp, Player, Nation, Pos, Age, MP_Playing, Starts_Playing, Min_Playing, Gls, Ast, PK, PKatt, xG_Per, npxG_Per, xA_Per) %>% 
  mutate(npxG_plus_xA_per = npxG_Per + xA_Per, 
         npxG_plus_xA_per_min = round(npxG_plus_xA_per * Min_Playing / max(Min_Playing), digits = 3)) %>% 
  arrange(-npxG_plus_xA_per_min) %>% 
  separate(Player, into = c("first_name", "last_name"), sep = " ")

```

Lets take a look the FPL 2022/2023 player list 
```{r}
ff <- read.csv("./00_raw_data/fpl_players.csv")

#AscToChar("<e9>")
#strsplit(AscToChar(ff$Player[4]), split=NULL)

ff <- ff %>% 
  left_join(fb, by=c("Player"="last_name")) %>% 
  drop_na() %>% 
  filter(Min_Playing > 10) %>% 
  mutate(residuals = round(residuals(loess(Draft.Rank ~ npxG_plus_xA_per_min)), digits = 2)) %>% 
  arrange(-npxG_plus_xA_per_min)

ff <- ff %>% 
  select(Player, Team, Age, Pos.x, Draft.Rank, npxG_plus_xA_per_min, residuals)

names(ff)[4] <- "Pos"

ff_for <- ff %>% 
  filter(Pos == "F")

ff_mid <- ff %>% 
  filter(Pos == "M")

ff_def <- ff %>% 
  filter(Pos == "D")

ff_gk <- ff %>% 
  filter(Pos == "GK")

```
Be warned, drop_na is removing about 200 players!

```{r}
ff_fpl_players <- fpl_player %>% 
  left_join(ff, by=c("web_name"="Player")) 

ff_forwards <- ff_fpl_players %>% 
  filter(Pos == "F") %>% 
  select(name, web_name, Team, Age, Draft.Rank, bps, mean_next_gw_points, points_per_game, influence, npxG_plus_xA_per_min, actual_vs_expected) %>% 
  unique()

ff_mids <- ff_fpl_players %>% 
  filter(Pos == "M") %>% 
  select(name, web_name, Team, Age, Draft.Rank, bps, mean_next_gw_points, points_per_game, influence, npxG_plus_xA_per_min, actual_vs_expected, corners_and_indirect_freekicks_order) %>% 
  unique()

ff_defenders <- ff_fpl_players %>% 
  filter(Pos == "D") %>% 
  select(name, web_name, Team, Age, Draft.Rank, bps, minutes, mean_next_gw_points, points_per_game, influence, npxG_plus_xA_per_min, actual_vs_expected, clean_sheets, corners_and_indirect_freekicks_order) %>% 
  unique()

ff_goalies <- ff_fpl_players %>% 
  filter(Pos == "GK") %>% 
  select(name, web_name, Team, Age, Draft.Rank,bps, value_season, clean_sheets, saves) %>% 
  unique()

```


Now lets sort the data
```{r}
test %>% 
  select(Player, Team, Draft.Rank, Points, npxG_plus_xA_per_min) %>% 
  arrange(-npxG_plus_xA_per_min) %>% 
  head(15) %>% 
  gt(rowname_col = "Player") %>% 
  tab_header(
    title = "Top 15 2021/2022 npxG_plus_xA_per_min") %>% 
  tab_source_note(
    source_note = "") %>% 
  tab_stubhead(label = "name") %>% 
  tab_source_note(source_note = "Twitter: @its_MikeF | Data from FBref") 
```

Lets plot the data
```{r}
{test %>%
    ggplot(aes(x = Draft.Rank, y = npxG_plus_xA_per_min)) +
    geom_smooth(method=loess, se=F) +
    geom_point(alpha = 0.7, cex = 3) +
    geom_text_repel(aes(label=Player)) +
    labs(x = "Draft.Rank",
         y = "npxG_plus_xA_per_min",
         title = "2022/2023 Value Finder",
         caption = "Twitter: Its_MikeF | Data: FBRef") +
    theme_bw() +
    theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold")) +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    scale_x_continuous(breaks = scales::pretty_breaks(n = 10))
  }
```

```{r}
fb_big5_advanced_season_stats_defense <- fb_big5_advanced_season_stats(
season_end_year = 2022,
stat_type = "defense",
team_or_player = "player"
)
```
stat_name = c("Accurate long balls per 90","Big chances created", "Clean sheets", "Clearances per 90","FotMob rating","Goals prevented", "xG + xA per 90")
