---
title: "soccer_learn"
output: html_document
date: '2022-07-25'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lets test out some soccer packages!

First we load in the packages
```{r}
suppressMessages({
  library(tidyverse) #metapackage
  library(gt) #Easily Create Presentation-Ready Display Tables
  library(ggsci) #Scientific Journal and Sci-Fi Themed Color Palettes for 'ggplot2'
  library(reshape2) #flexibly Reshape Data
  library(ggrepel) #Automatically Position Non-Overlapping Text Labels with 'ggplot2'
  library(webshot)
  library(DescTools) # Tools for Descriptive Statistics
  
  library(worldfootballR) #extract and clean world soccer data
  library(engsoccerdata) #English and European Soccer Results
  library(socceR) #Evaluating Sport Tournament Predictions
  library(ggsoccer) #Plot Soccer Event Data
  library(footBayes) #Fitting Bayesian and MLE Football Models
  library(itscalledsoccer) #American Soccer Analysis API Client
  library(FPLdata) #Read in Fantasy Premier League Data
  library(EUfootball) #Football Match Data of European Leagues
})
```

Lets check out FPL dataset
```{r}
fpl_data <- FPLdata()
names(fpl_data)
```
Lets now sort for player level season long data
```{r}
fpl_player <- fpl_data %>%
  unite(name ,first_name, second_name, sep = " ") %>% 
  group_by(name, web_name) %>%
  summarise(
    "minutes" = max(minutes),
    "chance_of_playing_this_round" = round(mean(chance_of_playing_this_round, na.rm = T), digits = 2),
    "mean_next_gw_points" = round(mean(next_gw_points, na.rm = T), digits = 2), 
    "bps" = max(bps, na.rm = T), 
    
    "form" = round(mean(form, na.rm = T), digits = 2),
    "value_form" = round(mean(value_form, na.rm = T), digits = 2),
    "value_season" = max(value_season, na.rm = T),
    
    "dream team" = max(dreamteam_count),
    "points_per_game" = round(mean(points_per_game, na.rm = T), digits =2), 
    
    "creativity" = round(mean(creativity, na.rm = T), digits =0),
    "creativity_rank" = round(min(creativity_rank, na.rm = T), digits =0),
    
    "influence" = round(mean(influence, na.rm = T), digits =0),
    "influence_rank" = round(min(influence_rank, na.rm = T), digits =0),
    
    "threat" = round(mean(threat, na.rm = T), digits =0),
    "threat_rank" = round(min(threat_rank, na.rm = T), digits =0),
    
    "clean_sheets" = round(sum(clean_sheets, na.rm = T), digits =0), 
    "value_season" = max(value_season, na.rm = T), 
    
    "saves" = max(saves, na.rm = T), 
    "corners_and_indirect_freekicks_order" = round(mean(corners_and_indirect_freekicks_order, na.rm = T),digits = 1)) %>%
  #filter(minutes > 0.25 * max(minutes)) %>% 
  mutate(actual_vs_expected = points_per_game-mean_next_gw_points) %>% 
  arrange(-actual_vs_expected) 
head(fpl_player)
```

Lets take a look to see which players over performed the most
```{r}
fpl_player %>% 
  select(name, minutes, mean_next_gw_points, points_per_game, actual_vs_expected) %>% 
  head(15) %>% 
  gt(rowname_col = "name") %>% 
  tab_header(
    title = "Top 15 2021/2022 FPL Overperformers", 
    subtitle = "Sorted by actual_vs_expected, filtered out minutes <20%") %>% 
  tab_source_note(
    source_note = "next_gw_points (next gameweek points) value for each player given their attributes in the current week.") %>% 
  tab_stubhead(label = "name") %>% 
  tab_source_note(source_note = "Twitter: @its_MikeF | Data from FPLdata") 
```

Lets see if threat or activity rank correlated
```{r}
normalize <- function(x){
  return((x - min(x,na.rm = T))/( max(x, na.rm = T) - min(x, na.rm = T)))
}

fpl_player_norm <- fpl_player

fpl_player_norm <- fpl_player_norm[-c(1,2,4,20)]

for(i in 1:length(fpl_player_norm)){
  fpl_player_norm[,i] = normalize(fpl_player_norm[,i])
}

fpl_player_norm %>%
  cor() %>%
  melt() %>%
  ggplot(aes(Var1, Var2, fill=value)) +
  geom_tile(color='white') +
  scale_fill_distiller(palette = 'GnBu', direction = 1) +
  geom_text(aes(label=paste(round(value,2)*100,'%')), size=2.5, color='black') +
  labs(x='',y='',fill='correlations', title='Relationship between fpl variables') +
  theme(axis.text.x = element_text(angle = 90, vjust = .5))
```
Lets check out the **worldfootballr** package
```{r}
fb_big5_advanced_season_stats <- fb_big5_advanced_season_stats(
season_end_year = 2022,
stat_type = "standard",
team_or_player = "player",
time_pause = 3
)
```

The statistic type options (stat_type) include:
"standard", "shooting", "passing", "passing_types", "gca", "defense", "possession", "playing_time", "misc", "keepers", "keepers_adv"

```{r}
fb <- fb_big5_advanced_season_stats %>% 
  select(Squad, Comp, Player, Nation, Pos, Age, MP_Playing, Starts_Playing, Min_Playing, Gls, Ast, PK, PKatt, xG_Per, npxG_Per, xA_Per) %>% 
  mutate(npxG_plus_xA_per = npxG_Per + xA_Per, 
         npxG_plus_xA_per_min = round(npxG_plus_xA_per * Min_Playing / max(Min_Playing), digits = 3)) %>% 
  arrange(-npxG_plus_xA_per_min) %>% 
  separate(Player, into = c("first_name", "last_name"), sep = " ")

```
Lets take a look the FPL 2022/2023 player list 
```{r}
ff <- read.csv("./00_raw_data/fpl_players.csv")

#AscToChar("<e9>")
#strsplit(AscToChar(ff$Player[4]), split=NULL)

ff <- ff %>% 
  left_join(fb, by=c("Player"="last_name")) %>% 
  drop_na() %>% 
  filter(Min_Playing > 10) %>% 
  mutate(residuals = round(residuals(loess(Draft.Rank ~ npxG_plus_xA_per_min)), digits = 2)) %>% 
  arrange(-npxG_plus_xA_per_min)

ff <- ff %>% 
  select(Player, Team, Age, Pos.x, Draft.Rank, npxG_plus_xA_per_min, residuals)

names(ff)[4] <- "Pos"

ff_for <- ff %>% 
  filter(Pos == "F")

ff_mid <- ff %>% 
  filter(Pos == "M")

ff_def <- ff %>% 
  filter(Pos == "D")

ff_gk <- ff %>% 
  filter(Pos == "GK")

```
Be warned, drop_na is removing about 200 players!

```{r}
ff_fpl_players <- fpl_player %>% 
  left_join(ff, by=c("web_name"="Player")) 

ff_forwards <- ff_fpl_players %>% 
  filter(Pos == "F") %>% 
  select(name, web_name, Team, Age, Draft.Rank, bps, mean_next_gw_points, points_per_game, influence, npxG_plus_xA_per_min, actual_vs_expected) %>% 
  unique()

ff_mids <- ff_fpl_players %>% 
  filter(Pos == "M") %>% 
  select(name, web_name, Team, Age, Draft.Rank, bps, mean_next_gw_points, points_per_game, influence, npxG_plus_xA_per_min, actual_vs_expected, corners_and_indirect_freekicks_order) %>% 
  unique()

ff_defenders <- ff_fpl_players %>% 
  filter(Pos == "D") %>% 
  select(name, web_name, Team, Age, Draft.Rank, bps, minutes, mean_next_gw_points, points_per_game, influence, npxG_plus_xA_per_min, actual_vs_expected, clean_sheets, corners_and_indirect_freekicks_order) %>% 
  unique()

ff_goalies <- ff_fpl_players %>% 
  filter(Pos == "GK") %>% 
  select(name, web_name, Team, Age, Draft.Rank,bps, value_season, clean_sheets, saves) %>% 
  unique()

```


Now lets sort the data
```{r}
test %>% 
  select(Player, Team, Draft.Rank, Points, npxG_plus_xA_per_min) %>% 
  arrange(-npxG_plus_xA_per_min) %>% 
  head(15) %>% 
  gt(rowname_col = "Player") %>% 
  tab_header(
    title = "Top 15 2021/2022 npxG_plus_xA_per_min") %>% 
  tab_source_note(
    source_note = "") %>% 
  tab_stubhead(label = "name") %>% 
  tab_source_note(source_note = "Twitter: @its_MikeF | Data from FBref") 
```

Lets plot the data
```{r}
{test %>%
    ggplot(aes(x = Draft.Rank, y = npxG_plus_xA_per_min)) +
    geom_smooth(method=loess, se=F) +
    geom_point(alpha = 0.7, cex = 3) +
    geom_text_repel(aes(label=Player)) +
    labs(x = "Draft.Rank",
         y = "npxG_plus_xA_per_min",
         title = "2022/2023 Value Finder",
         caption = "Twitter: Its_MikeF | Data: FBRef") +
    theme_bw() +
    theme(plot.title = element_text(size = 14, hjust = 0.5, face = "bold")) +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    scale_x_continuous(breaks = scales::pretty_breaks(n = 10))
  }
```

```{r}
fb_big5_advanced_season_stats_defense <- fb_big5_advanced_season_stats(
season_end_year = 2022,
stat_type = "defense",
team_or_player = "player"
)


```
stat_name = c("Accurate long balls per 90","Big chances created", "Clean sheets", "Clearances per 90","FotMob rating","Goals prevented", "xG + xA per 90")
